---
title: "Kaasaegsed meetodid molekulaarses ökoloogias"
subtitle: "MLB7052.LT — Modern Molecular Methods in Ecology · 2026/2027 autumn"
toc: false
# Reads a CSV that the bots update; must not be served from the freeze cache.
execute:
  freeze: false
---

```{r}
#| label: setup
#| include: false
library(dplyr)
library(readr)
library(knitr)
source("R/course.R")

schedule <- read_csv("data/schedule.csv", show_col_types = FALSE)

# Every date on the site is derived from the single semester-start date in
# _quarto.yml, so shifting the semester is a one-line change.
start   <- as.Date(course_meta("semester-start"))
schedule <- schedule |> mutate(date = start + 7 * (week - 1))
```

From the chemistry of a nucleic acid to a defensible ecological inference. The
course follows one chain of reasoning all semester:

::: {style="text-align:center; font-size:1.05em; margin:1.2em 0; padding:0.9em; background:#f7f3f3; border-left:4px solid #990000;"}
**biological question → sample → molecule → assay → sequence or measurement → computation → inference**
:::

[🗓️ Schedule](schedule.qmd){.cta}
[👨‍🏫 Lectures](#lectures){.cta}
[🎤 Register your seminar](seminars/index.qmd){.cta}
[✅ Log your attendance](participation.qmd){.cta .ghost}
[📓 Syllabus](syllabus.qmd){.cta .ghost}

::: {.callout-important}
## Two things to do in week 1

1. Send me your **GitHub username** so you can be added to this repository.
2. Read [how seminars work](seminars/index.qmd) and start looking for a paper.
   Registration is first-come, first-served.
:::

# What the course is about

Molecular data have become the default evidence in ecology: a water sample
stands in for a fish survey, a gut content becomes a diet table, a handful of
sediment becomes a century of community change. Every one of those inferences
depends on choices made long before the sequencer — what was sampled, how it was
preserved, which molecule was targeted, which controls were run.

This course teaches the methods *and* the reasoning that makes their output
trustworthy. You will learn what each technique measures, what it cannot
measure, and how to tell the difference in someone else's paper.

By the end you should be able to design a molecular study end-to-end, run the
core bench and computational steps, and defend the limits of what your data
support. The four content blocks are laid out on the [topics page](topics.qmd),
and the week-by-week plan on the [schedule](schedule.qmd).

# Lectures {#lectures}

```{r}
#| label: lecture-cards
#| echo: false
#| results: asis

cards <- schedule |> filter(!is.na(lecture), lecture != "")

cat('<div class="lecture-grid">\n')
for (i in seq_len(nrow(cards))) {
  r <- cards[i, ]
  cat(sprintf(
    '<div class="lecture-card"><a href="%s"><h4>%s</h4></a><p>Week %d · %s · Block %s</p></div>\n',
    r$lecture,
    htmltools::htmlEscape(r$topic),
    r$week,
    format(r$date, "%d %b %Y"),
    r$block
  ))
}
cat('</div>\n')
```

Slides open in your browser; use the arrow keys, `F` for fullscreen, `S` for
speaker notes and `?` for the full key list. Remaining weeks are listed on the
[schedule](schedule.qmd) and appear here as they are delivered.

# Practical information

```{r}
#| label: course-info
#| echo: false
meta <- course_meta()
tibble::tribble(
  ~Field,          ~Detail,
  "Course code",   meta[["course-code"]],
  "Title",         paste0(meta[["course-title-et"]], " / ", meta[["course-title-en"]]),
  "Semester",      meta[["semester"]],
  "Teaching time", "Fridays, weekly — see the schedule for dates",
  "Language",      "Estonian teaching; English reading and software documentation",
  "Prerequisites", "MLK6922.LT Biochemistry; MLB6013.LT Genetics",
  "Lecturer",      "Peeter Laas, PhD — peeter.laas@ut.ee",
  "Assessment",    "Graded A–F; see the syllabus for the component weights"
) |> knitr::kable()
```

# How this site works

Everything runs through this repository, so there is one place to look and one
place to submit.

| You want to… | Where | How |
|---|---|---|
| See what is coming | [Schedule](schedule.qmd) | Week-by-week plan with links to slides |
| Read the slides | [Lectures](#lectures) | Rendered in the browser, no download needed |
| Record that you attended | [Participation](participation.qmd) | Open an issue from the attendance form |
| Claim a paper for your seminar | [Seminars](seminars/index.qmd) | Open an issue; a bot adds you to the table |
| Share your seminar slides | [Seminars](seminars/index.qmd) | Pull request with a PDF, PPTX or QMD |
| Find a reading or a tool | [Materials](materials.qmd) | Books, software, databases, datasets |

No Google Forms, no separate sign-ups: your GitHub account is your identity for
attendance and for seminar registration, and the record lives in this
repository where you can check it yourself.
