---
title: "✅ Participation"
subtitle: "Attendance is recorded through this repository — no forms, no sign-ups"
# Reads a CSV that the bots update; must not be served from the freeze cache.
execute:
  freeze: false
---

```{r}
#| label: setup
#| include: false
library(dplyr)
library(tidyr)
library(readr)
library(knitr)
source("R/course.R")

repo       <- course_meta("course-repo")
sessions   <- read_csv("data/sessions.csv", show_col_types = FALSE,
                       col_types = cols(.default = col_character()))
attendance <- read_csv("data/attendance.csv", show_col_types = FALSE,
                       col_types = cols(.default = col_character()))

open_sessions <- sessions |> filter(tolower(open) == "yes")
new_issue     <- paste0(repo, "/issues/new?template=attendance.yml")
```

# How to log a session

```{r}
#| label: open-now
#| echo: false
#| results: asis

if (nrow(open_sessions) > 0) {
  cat('::: {.callout-important appearance="simple"}\n## Open right now\n\n')
  for (i in seq_len(nrow(open_sessions))) {
    s <- open_sessions[i, ]
    cat(sprintf("- **%s** — %s (%s)\n", s$session_id, s$topic, s$date))
  }
  cat(sprintf('\n[Log your attendance](%s){.cta}\n\n:::\n', new_issue))
} else {
  cat('::: {.callout-note appearance="simple"}\n## Nothing open at the moment\n\n')
  cat('No session is currently accepting attendance. The form opens during class ')
  cat('and closes shortly after it ends.\n\n:::\n')
}
```

1. During the session, open the attendance form:
   [**Log attendance**](`r new_issue`).
2. Pick the session and enter the **code given in class**.
3. Submit. A bot checks the code, adds you to `data/attendance.csv`, and closes
   the issue with a ✅ or an explanation of what went wrong.

That is the whole process. Your GitHub username is your identity, so there is
nothing to type but the session and the code, and nothing to lose track of: the
record is a file in this repository that you can read yourself at any time.

::: {.callout-note collapse="true"}
## Why it works this way

An earlier plan used a Google Form feeding a spreadsheet. Doing it in GitHub
instead means one account instead of two, no personal data leaving the
university's control, no email addresses collected at all, and an audit trail —
every attendance record is a commit made by a bot in response to an issue you
opened, and both halves stay visible.

The session code exists so that attendance means *being in the room*. Codes are
announced in class and are not published here.
:::

# The record

```{r}
#| label: attendance-matrix
#| echo: false
#| results: asis

if (nrow(attendance) == 0) {
  cat("No sessions have been logged yet. The table appears here after the first one.\n")
} else {
  seen <- sessions |> filter(session_id %in% attendance$session_id)

  grid <- attendance |>
    distinct(github_user, session_id) |>
    mutate(present = "✓") |>
    complete(github_user, session_id = seen$session_id, fill = list(present = "·")) |>
    pivot_wider(names_from = session_id, values_from = present) |>
    arrange(github_user)

  totals <- attendance |>
    distinct(github_user, session_id) |>
    count(github_user, name = "Total")

  out <- grid |>
    left_join(totals, by = "github_user") |>
    rename(`GitHub user` = github_user)

  print(kable(out, align = c("l", rep("c", ncol(out) - 1))))

  cat(sprintf("\n\n%d session%s logged by %d student%s.\n",
              length(unique(attendance$session_id)),
              if (length(unique(attendance$session_id)) == 1) "" else "s",
              length(unique(attendance$github_user)),
              if (length(unique(attendance$github_user)) == 1) "" else "s"))
}
```

# Sessions

```{r}
#| label: session-table
#| echo: false
counts <- attendance |> count(session_id, name = "Logged")

sessions |>
  left_join(counts, by = "session_id") |>
  transmute(
    Session = session_id,
    Date    = date,
    Topic   = topic,
    Status  = ifelse(tolower(open) == "yes", "open", "closed"),
    Logged  = ifelse(is.na(Logged), 0L, as.integer(Logged))
  ) |>
  kable()
```

# If something is wrong

Check `data/attendance.csv` first — it is the record, and it is readable. If a
session you attended is missing, or one you did not attend is present, open a
plain issue describing the problem. Do not edit the file directly: the bot owns
it, and a manual edit will be overwritten.

Corrections are made by the lecturer as a normal commit, so the change and its
reason stay visible.
