---
title: "🗓️ Schedule"
subtitle: "Week-by-week plan for the 2026/2027 autumn semester"
# Reads a CSV that the bots update; must not be served from the freeze cache.
execute:
  freeze: false
---

```{r}
#| label: setup
#| include: false
library(dplyr)
library(readr)
library(knitr)
source("R/course.R")

schedule <- read_csv("data/schedule.csv", show_col_types = FALSE)
start    <- as.Date(course_meta("semester-start"))

schedule <- schedule |>
  mutate(
    date  = start + 7 * (week - 1),
    past  = date < Sys.Date(),
    # A row links to its deck when one exists, otherwise the topic is plain text.
    topic_cell = ifelse(
      is.na(lecture) | lecture == "",
      topic,
      sprintf("[%s](%s)", topic, lecture)
    ),
    state = ifelse(past, "delivered", status)
  )
```

::: {.callout-tip appearance="simple"}
Dates are derived from a single semester-start date in `_quarto.yml`
(**`r format(start, "%d %B %Y")`**, the first teaching Friday). Confirm it
against the official timetable — changing that one line moves the whole
schedule.
:::

```{r}
#| label: schedule-table
#| echo: false
schedule |>
  transmute(
    Week    = week,
    Date    = format(date, "%a %d %b"),
    Block   = block,
    Format  = kind,
    Topic   = topic_cell,
    Reading = ifelse(is.na(reading) | reading == "", "—", reading)
  ) |>
  kable()
```

# Blocks

The semester is organised into four blocks; the
[topics page](topics.qmd) lists what each one covers.

```{r}
#| label: block-summary
#| echo: false
block_names <- c(
  I   = "Foundations and measurement",
  II  = "The molecular toolkit",
  III = "Genomes and molecular phenotypes",
  IV  = "Ecology, evolution, and application"
)

schedule |>
  group_by(block) |>
  summarise(
    Weeks = paste0(min(week), "–", max(week)),
    Sessions = n(),
    .groups = "drop"
  ) |>
  transmute(
    Block = paste0("Block ", block),
    Title = block_names[block],
    Weeks,
    Sessions
  ) |>
  kable()
```

# What happens in a typical week

| | |
|---|---|
| **Before** | Read the assigned chapter or paper; the reading quiz opens with it |
| **In class** | Lecture, then structured interpretation of one figure, dataset or method decision |
| **After** | Log your [attendance](participation.qmd); write up any practical from that week |

Weeks 7 and 15 differ: week 7 is a full practical session, and week 15 is given
over to [student seminar presentations](seminars/index.qmd).

# Changes

The schedule lives in `data/schedule.csv`. Any change to it is a commit, so the
history of what moved and when is visible in the repository — there is no
separate announcement channel to miss.
