---
title: "🎤 Seminars"
subtitle: "Claim a paper, present it, share your slides"
# Reads a CSV that the bots update; must not be served from the freeze cache.
execute:
  freeze: false
---

```{r}
#| label: setup
#| include: false
library(dplyr)
library(readr)
library(knitr)
source("../R/course.R")

repo  <- course_meta("course-repo")
regs  <- read_csv("registrations.csv", show_col_types = FALSE,
                  col_types = cols(.default = col_character()))

claim_issue <- paste0(repo, "/issues/new?template=seminar.yml")
slides_dir  <- "slides"

# A registration is "delivered" once a matching file has landed in slides/.
have_file <- function(f) !is.na(f) && nzchar(f) && file.exists(file.path(slides_dir, f))
```

Each of you presents **one research article**, 15 minutes plus 3–5 for
questions, in English. It is worth **20%** of the course grade. Pick a paper
that argues about *how evidence was produced*, not only about what was found —
the seminar is a methods seminar.

[📝 Claim a paper](`r claim_issue`){.cta}
[📤 Submit your slides](#submitting-your-slides){.cta .ghost}

# 1. Claim a paper

Open the registration form and give the title and the DOI or URL. A bot checks
that nobody has claimed that paper already, adds a row to
`registrations.csv`, and closes the issue confirming it. **First come, first
served** — if two people claim the same paper, the earlier issue wins and the
second is told to choose again.

Before you claim, check the table below and skim the journals suggested on the
[materials page](../materials.qmd).

# 2. The register

```{r}
#| label: register
#| echo: false
#| results: asis

if (nrow(regs) == 0) {
  cat("Nothing claimed yet — the table fills up as registrations come in.\n")
} else {
  regs |>
    mutate(
      Paper = ifelse(
        is.na(paper_doi_or_url) | paper_doi_or_url == "",
        paper_title,
        sprintf("[%s](%s)", paper_title, paper_doi_or_url)
      ),
      Slides = vapply(slides_file, function(f) {
        if (have_file(f)) sprintf("[⬇ %s](%s/%s)", f, slides_dir, f) else "—"
      }, character(1))
    ) |>
    transmute(
      Student = ifelse(is.na(student_name) | student_name == "",
                       github_user, student_name),
      `GitHub` = paste0("@", github_user),
      Paper,
      Date = ifelse(is.na(presentation_date) | presentation_date == "",
                    "to be scheduled", presentation_date),
      Slides
    ) |>
    kable() |>
    print()
}
```

# 3. Submitting your slides {#submitting-your-slides}

Slides go into `seminars/slides/` in this repository, as a **pull request**.
Any of three formats is fine:

| Format | Extension | Note |
|---|---|---|
| PDF | `.pdf` | Safest — renders identically everywhere |
| PowerPoint | `.pptx` | Fine; export a PDF too if you used unusual fonts |
| Quarto | `.qmd` | Welcome, and the most reproducible. It is linked for download, not rendered by the site |

Name the file `githubusername-shortpapertitle.pdf` — for example
`jsmith-edna-detection-limits.pdf`.

### With the GitHub website

1. Go to [`seminars/slides/`](`r paste0(repo, "/tree/main/seminars/slides")`) in the repository.
2. **Add file → Upload files**, drag your slides in.
3. Choose **Create a new branch and start a pull request**, then **Propose changes**.
4. Also add the file name to your row's `slides_file` column in
   `registrations.csv`, in the same pull request.

### With git

```bash
git clone https://github.com/peeterlaas/MolEcoMeth.git
cd MolEcoMeth
git checkout -b slides-yourname
cp ~/Desktop/yourname-yourpaper.pdf seminars/slides/
# add the file name to your row in seminars/registrations.csv
git add seminars/
git commit -m "Add seminar slides for <paper>"
git push -u origin slides-yourname
```

Then open the pull request from the link git prints.

A pull request that only touches `seminars/` is checked and merged
automatically. One that touches anything else waits for the lecturer — that is
deliberate, and it is the reason you can be given write access to the repository
at all.

::: {.callout-important}
## Deadline

Slides are due **before** your presentation slot. Submit the version you will
present; if you fix a typo afterwards, open a second pull request rather than
rewriting history.
:::

# What makes a good seminar

You are not summarising the paper. You are explaining a method well enough that
the room can judge it.

- **The question** — what did they want to know, in one sentence?
- **The evidence chain** — sample → molecule → assay → data → claim. Walk it.
- **The controls** — what did they run, and what would have failed without them?
- **The weak joint** — every study has one step where the inference gets
  fragile. Name it, and say what you would have done instead.
- **The transfer** — where else does this method apply, and where would it break?

Slides should be readable from the back of the room: few words, one idea per
slide, and the paper's key figure large enough to actually interpret.

# Grading

| | Weight |
|---|---:|
| Understanding of the method and its limits | 40% |
| Clarity of the evidence chain | 25% |
| Quality of the slides and delivery | 20% |
| Handling of questions | 15% |
